On Verification Patterns for Low-Level Systems via Modal Abstractions

### 2025-09-11 paper.tex
sigpln
+

%%
%% This is file `sample-sigconf-authordraft.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-authordraft.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,10pt]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.

  In this paper we examine how modalities have emerged as a common structure in formal verification
  of low-level software, and explain how many recent examples naturally share 
  common structure in the relationship between the modalities and software features they are used
  to reason about.
  We explain how the concept of a \emph{resource context} (a class of system resources
  to reason about) naturally corresponds to families of modal operators indexed by
  system data, and how this naturally leads to using modal assertions to describe
  \emph{resource elements} (data in the relevant context).
  We also describe extensions of this idea to ongoing work in formal verification
  of filesystems code.
  % \todo[inline,color=cyan]{Ismail, why did you comment out my suggestion and keep the original?}
  % \todo[inline,color=pink]{Sorry, I thought I did the other way}
\end{abstract}

@@

% \subsection{Conclusion \& Contributions}
% \paragraph{Conclusion} 
Low-level systems software exhibits certain patterns in their designs, especially when interacting with computing resources. 
Exploiting certain patterns while designing software has been an important field of study.
This survey examines these common patterns in systems software, and their relationships
to modalities. We believe that certain properties of modalities enable us to understand and 
address the verification challenges of these systems by tailoring specification and proof
to the same patterns used to design the systems.

\textit{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges. We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
\begin{itemize}
\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource}, which has already been an essential concept in the design of systems. Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
\item \textbf{Introducing \emph{Nominals for System Resources}}:
Existing logics for systems software must frequently identify
individual resources within a given context ---
nominalization (in the sense of hybrid logic) enables identifying a resource in its context. 
For example, a filesystem transaction is a context of resources for in-memory updated disk blocks
which have not yet been saved to disk.
The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, 
in case of a crash while persisting updated disk-blocks, the filesystem can rollback the 
\emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state. 
We discuss other examples to contextualize the existing systems surveyed.
% to be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
% identifier -- \emph{strong nominalization}. 
% in another example, virtual memory references (resources) in 
% an address space (resource context), which can be uniquely identified with a root address (the nominal) 
% of its page-table tree, are \emph{agnostic} to the address space that they are in. 
% However, they can only be accessed (be valid) in the address space to which they are agnostic. 
% However, then an address space switch happens, the virtual memory references of the previous address 
% space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
% piece of information related to their address space, we still have to associate them. 
% We call this kind of unilateral nominalization -- \emph{weak nominalization}.
% \todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
\item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions. We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
\end{itemize}

>>>>

%%
%% This is file `sample-sigconf-authordraft.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,authordraft')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-authordraft.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigplan,10pt]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Although they differ in the functionality they offer, low-level systems exhibit certain patterns of design and utilization of computing resources.
  In this paper we examine how modalities have emerged as a common structure in formal verification
  of low-level software, and explain how many recent examples naturally share 
  common structure in the relationship between the modalities and software features they are used
  to reason about.
  We explain how the concept of a \emph{resource context} (a class of system resources
  to reason about) naturally corresponds to families of modal operators indexed by
  system data, and how this naturally leads to using modal assertions to describe
  \emph{resource elements} (data in the relevant context).
  We also describe extensions of this idea to ongoing work in formal verification
  of filesystems code.
  % \todo[inline,color=cyan]{Ismail, why did you comment out my suggestion and keep the original?}
  % \todo[inline,color=pink]{Sorry, I thought I did the other way}
\end{abstract}

@@

% \subsection{Conclusion \& Contributions}
% \paragraph{Conclusion} 
Interacting with computing resources is the essential point which shapes the design of low-level systems software. 
Exploiting these design choices has been an important field of study.
This survey examines these common patterns in systems software, and their relationships
to modalities. We believe that certain properties of modalities enable us to understand and 
address the verification challenges of these systems by tailoring specification and proof
to the same patterns used to design the systems.

\textit{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges. We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
\begin{itemize}
\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems. Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
\item \textbf{Introducing \emph{Nominals for System Resources}}:
Existing logics for systems software must frequently identify
individual resources within a given context ---
nominalization (in the sense of hybrid logic) enables identifying a resource in its context. 
For example, a filesystem transaction is a context of resources for in-memory updated disk blocks
which have not yet been saved to disk.
The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, 
in case of a crash while persisting updated disk-blocks, the filesystem can rollback the 
\emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state. 
We discuss other examples to contextualize the existing systems surveyed.
% to be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
% identifier -- \emph{strong nominalization}. 
% in another example, virtual memory references (resources) in 
% an address space (resource context), which can be uniquely identified with a root address (the nominal) 
% of its page-table tree, are \emph{agnostic} to the address space that they are in. 
% However, they can only be accessed (be valid) in the address space to which they are agnostic. 
% However, then an address space switch happens, the virtual memory references of the previous address 
% space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
% piece of information related to their address space, we still have to associate them. 
% We call this kind of unilateral nominalization -- \emph{weak nominalization}.
% \todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
\item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions. We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
\end{itemize}
